### 2024-12-01 paper/paper.qmd
Upload for Introduction paragraph

---
title: "Pricing Patterns in Organic Markets"
subtitle: " Comparative Analysis of Vendor Strategies and Seasonal Trends in Organic Product Sales"
author: 
  - Dingshuo Li
thanks: "Code and data are available at: [https://github.com/dawsonlll/Organic-Product-Pricing-Analysis.git](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format:
  pdf:
    toc: true
  html:
    toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(rstanarm)
library(modelsummary)
library(ggplot2)
library(arrow)
library(here)
library(modelr)
library(knitr)
library(rstanarm)
library(broom)
library(scales)
library(splines)
opts_knit$set(root.dir = here::here())
```

```{r}
#| include: false
#| warning: false
#| message: false
setwd(here::here())
organic_analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
bayesian_price_model <- readRDS(file = here::here("models/bayesian_price_model.rds"))
```

# Introduction

The organic food market has witnessed a remarkable surge in demand, driven by increasing consumer awareness of health and environmental impacts. This demand has not only heightened the market value of organic products but has also introduced complexities into their pricing strategies across various retail platforms. Notably, organic food typically incurs higher production and certification costs compared to conventional food products, influencing its retail pricing and accessibility. This study aims to dissect these pricing strategies by examining how the prices of organic foods vary across different vendors and over successive months.

This paper explores the intricate dynamics of organic food pricing within a well-defined dataset that records prices from multiple vendors over a specified period. The analysis is structured to address a gap in existing research: the detailed quantification of how prices for organic products are structured and vary in a real-market scenario. Utilizing advanced Bayesian statistical methods, this research assesses the relationship between current and historical prices while considering the influence of temporal factors captured through monthly variations.

Estimand: The primary estimand of this study is the price elasticity of organic foods across different market vendors and months. Specifically, the analysis seeks to estimate how changes in the historical prices (old prices) influence the current pricing strategies, adjusting for variations across months and between vendors. This will allow us to quantify the direct and interactive effects of time (month) and vendor on the current prices of organic products, providing insights into the responsiveness of organic food prices to market dynamics and vendor-specific factors.

The findings reveal distinct pricing patterns that differ by vendor and demonstrate notable fluctuations corresponding to seasonal trends. These insights are crucial for consumers making informed purchasing decisions and for vendors strategizing to capture or expand market share in the competitive organic food sector. The importance of understanding these pricing mechanisms lies in its potential to influence marketing strategies, consumer spending, and ultimately, the economic sustainability of organic farming practices.

The paper is structured to provide a clear, logical flow from methodology to implications @sec-data outlines our dataset selection and cleaning, ensuring transparency. The @sec-model details the Bayesian model and its rationale, directly linked to our data preparation. The @sec-result displays our findings with visual aids, while the @sec-discussion interprets these findings in the context of broader market dynamics. Finally, @sec-limitation  highlights areas for further study, acknowledging any constraints encountered.

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

```{r}

```

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
modelsummary::modelsummary(
  list(
    "Bayesian_model" = bayesian_price_model
  ),
  statistic = "mad",
  fmt = 2
)

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(bayesian_price_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(bayesian_price_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(bayesian_price_model, "trace")

plot(bayesian_price_model, "rhat")
```

\newpage

# References

### 2024-12-02 paper/paper.qmd
Paper update
Data, Model, Result finished, updated references

@@
 library(scales)
+library(bayesplot)
+
 library(splines)
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+All data analysis was conducted using R @citeR, including statistical computing and graphics.
+And the following R packages were used for conduct code in scripts: tidyverse @tidy, Palmerpenguins @palmerpenguins, ggplot2 @ggplot2, Dplyr @citedplyr, Knitr @citeknitr, modelsummary @modelsummary, arrow @arrow, here @here, rstanarm @rstanarm, sclaes @scales, splines @splines,broom @broom, lubridate @date, janitor @jan and testthat @testthat.
+Following  @tellingstories we conducted data simulate, test simulated data, data cleanning, test analysis data, EDA, data modelling.
+The R code in scripts were adapted from @alexander2023telling
 
-Overview text
+The dataset under analysis was sourced from Jacob Filipp's Hammer Project @hammerData, which compiles extensive pricing data on organic food products sold across various retail vendors.
+This dataset is particularly valuable for its comprehensive coverage of both temporal and vendor-specific pricing strategies, offering a unique lens through which seasonal trends and competitive behaviors in the organic food market can be observed.
+The choice of this dataset was motivated by its granularity and the quality of its data, which includes multiple vendors and a wide temporal range, making it ideal for a nuanced analysis of pricing dynamics.
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The dataset used in this study comes from Jacob Filipp's Hammer Project, which provides detailed records of organic food pricing across retailers and months.
+This data collection effort captures an important economic and social phenomenon: fluctuating pricing strategies in the competitive organic food market.
+Prices are recorded monthly to reflect the dynamic nature of these market conditions, providing a comprehensive snapshot that facilitates analysis of trends and pricing behavior over time.
+
+The primary step in capturing the phenomenon involves systematically collecting price data directly from retail sources.
+This data is gathered from grocery merchants.
+Each price tag observed represents a specific decision by a vendor regarding how much to charge for an organic product at a given time.
+So the important variables: current price which is the price of an organic product at the time of data collection.
+It is observed directly from the vendor.
+And old price which is the previously recorded price for the same product, which provides a historical comparison point.
+It helps in understanding how prices are adjusted over time.
+
+Each recorded price becomes an entry in the dataset, associated with metadata that includes the vendor, the time of collection, and the product details.
+This structured approach ensures that each entry in the dataset is a reliable and accurate representation of the real-world pricing phenomena observed.
+These entries are then ready for analytical processes that can derive meaningful insights about organic food pricing dynamics.
+
+## Data Cleanning
+The analysis begins with the consolidation of raw data from two distinct sources: product transaction records and product details.
+These datasets were sourced from the Hammer Project, containing over 12 million rows of diverse product data, including non-organic items and a variety of vendors beyond the scope of this study.
+The transaction records (raw_data) include timestamps, vendor details, and product identifiers, while the product details (product_data) provide comprehensive information such as product names, brands, pricing history, and unit measurements.
+
+To  conduct analysis, the two datasets were merged using an inner join on the product_id, which serves as a unique identifier linking transaction records to specific product details.
+This joining process ensured that each transaction record was enriched with relevant product information, creating a comprehensive view of each sale's context.
+
+Post-merger, the dataset still contained an extensive range of products and vendors, many of which were irrelevant to the focus of this study—organic food pricing dynamics among major vendors.
+To refine the dataset first did vendor selection.
+The dataset was filtered to include only transactions from major retailers known for their significant market presence and influence in organic food sales, specifically "Metro," "NoFrills," "Walmart," "Voila," "TandT," and "Galleria." This selection was based on the objective to analyze pricing strategies within the most competitive segments of the retail market.
+A further filtration was applied to isolate organic products.
+This was achieved by searching for the keyword "organic" within the product_name field, using a case-insensitive search to ensure comprehensive inclusion of all relevant items.
+This step significantly reduced the dataset size and increased its relevance to the study’s aims.
+
+The current_price and old_price were parsed to ensure that they were represented as numerical values, free from any non-numeric characters that could interfere with quantitative analysis.
+This parsing was crucial for accurate statistical calculations and comparisons.
+The nowtime field, representing the timestamp of each transaction, was used to extract the month component.
+This transformation was essential for analyzing seasonal trends and month-to-month pricing variations.
+Lastly, entries with missing or zero values in current_price were removed to maintain the integrity of pricing analysis.
+Additionally, any remaining rows with incomplete data were dropped, ensuring that the dataset used in subsequent analyses was entirely robust and clean.
 
 ## Outcome variables
-
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+### Predictor variables
+The old price variable, representing the product's price from the preceding month, serves as a critical metric for understanding the evolution of market trends and pricing strategies over time.
+The summary @tbl-old-price reveal a broad range of "Old Price" from a minimum of \$1.69 to a maximum of \$38.56, with a median price of \$5.49, suggesting a median market position.
+The mean, slightly higher at \$7.66, indicates a positive skew in the data, with the first quartile at \$3.99 and the third at $10.49, reflecting substantial variability in historical pricing.
+This variability underscores the diverse pricing strategies across different vendors and product types, and provides a foundation for analyzing how prices adjust in response to market dynamics, consumer demand, and possibly supply chain fluctuations.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| label: tbl-old-price
+#| tbl-cap: Summary of old price
 #| echo: false
@@
 #| message: false
+old_price_summary <- summary(organic_analysis_data$old_price)
+
+# Convert summary statistics into a data frame for better table formatting
+old_price_summary_df <- data.frame(
+  Statistic = c("Min", "1st Qu.", "Median", "Mean", "3rd Qu.", "Max"),
+  Value = as.numeric(old_price_summary)
+)
+
+# Print the summary as a table
+kable(old_price_summary_df, format = "markdown")
+```
+The vendor variable acts as a categorical identifier for retailers within the dataset, providing crucial insights into the diversity of pricing strategies employed across different retail environments.
+This variable is essential for comparing and contrasting market behaviours as it delineates the dataset into distinct groups based on the retailer.
+The distribution of data @tbl-Vendor shows that vendors varies significantly, with Metro being the most frequently occurring vendor with 1,981 entries, followed by Voila with 1,513 entries.
+Walmart, Galleria, TandT, and NoFrills appear with decreasing frequency, suggesting a wide range in the scale and possibly the geographic reach of these retailers.
+This spread indicates that Metro and Voila dominate this dataset, which could imply a greater influence or market share in the organic products sector compared to the other vendors.
+Understanding the pricing strategies of these vendors is key to analyzing market trends, as each vendor may react differently to market pressures, consumer preferences, and seasonal changes.
+Analyzing this variable can reveal patterns that are specific to certain types of retailers and provide valuable insights into the competitive landscape of the organic food market.
+```{r}
+#| label: tbl-Vendor
+#| tbl-cap: Preview of Breakdown of Organic Product Listings by Vendor
+#| echo: false
+#| warning: false
+#| message: false
+vendor_summary <- organic_analysis_data %>%
+  count(vendor) %>%
+  arrange(desc(n))
+
+# Print the table using knitr's kable function
+kable(vendor_summary, format = "markdown", col.names = c("Vendor", "Frequency"))
+
+```
+The "Month" variable in the dataset, recorded as a numerical value, serves as a crucial indicator for analyzing seasonal impacts on organic food pricing.
+It reflects when the data was collected, revealing essential patterns of how prices fluctuate through different months, likely influenced by consumer demand, supply variations, and promotional periods.
+The distribution of observations shown in @tbl-month, the peaks are in summer season suggests significant market activity during the summer months, potentially driven by an increased availability of fresh produce or higher consumer buying rates.
+In contrast, the notably lower entries in spring and winter indicate reduced activity, which could be due to offseason market slowness or less frequent data collection.
+This variable is integral to understanding how temporal factors influence pricing strategies, providing valuable insights for vendors, consumers, and analysts studying market dynamics in the organic food sector.
+```{r}
+#| label: tbl-month
+#| tbl-cap: Preview of Breakdown of Organic Product Listings by Month
+#| echo: false
+#| warning: false
+#| message: false
+month_summary <- organic_analysis_data %>%
+  count(month) %>%
+  arrange(month)
+kable(month_summary, format = "markdown", col.names = c("Month", "Count"))
 
@@
 
-Talk way more about it.
+### Response Variable
+The response variable current price represents the most recent price at which organic food products were sold across various vendors.
+The summary statistics @tbl-current-price for this variable provide a snapshot of its distribution within the dataset.
+The minimum price recorded is 0.97, indicating a lower-end pricing point for some products, while the maximum price peaks at 29.99, reflecting premium offerings within the organic sector.
+The first quartile is 3.22, suggesting that at least 25% of the products are priced at or below this level, which could be indicative of basic or smaller-sized organic goods.
+The median price of 4.49 and the mean of 5.69 suggest a central tendency towards moderately priced items, with the mean being slightly higher due to the influence of more expensive products pulling the average up.
+The third quartile at 7.49 shows that 75% of the products fall below this price point, highlighting a pricing structure that might cater more to a mid-range consumer segment.
+This comprehensive look at current price helps in understanding the economic landscape of organic food pricing and its variability, which is crucial for analyzing market strategies and consumer behaviour.
+```{r}
+#| label: tbl-current-price
+#| tbl-cap: Preview of Summary of Current Price
+#| echo: false
+#| warning: false
+#| message: false
+current_price_summary <- summary(organic_analysis_data$current_price)
 
-## Predictor variables
+# Convert summary statistics into a data frame for better table formatting
+current_price_summary_df <- data.frame(
+  Statistic = c("Min", "1st Qu.", "Median", "Mean", "3rd Qu.", "Max"),
+  Value = as.numeric(current_price_summary)
+)
+kable(current_price_summary_df, format = "markdown", col.names = c("Statistic", "Value"))
 
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+```
 
 # Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+## Model Overview
+In our study, we are particularly interested in understanding how the current price of organic products is influenced by several key factors: the price of the product in the previous month (old price), the vendor from which the product is purchased, and the time of year (month).
+To systematically examine these relationships, we developed a Bayesian regression model.
+This model incorporates these variables to elucidate their impacts on current pricing strategies.
+By leveraging historical pricing data alongside vendor-specific information and temporal dynamics, our model aims to uncover meaningful patterns that can inform both theoretical understanding and practical applications in market pricing strategies.
 
 ## Model set-up
+```{=latex}
+\[
+\text{Current Price}_i = \beta_0 + \beta_1 \times \text{Month}_i + \beta_2 \times \text{Old Price}_i + \sum (\beta_{\text{vendor}} \times \text{Vendor}_i) + \epsilon_i
+\]
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+\text{Where:}
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-```{r}
+\begin{itemize}
+  \item $\text{Current Price}_i$ is the dependent variable representing the price of organic products at time $i$.
+  \item $\text{Month}_i$ is a continuous independent variable reflecting the month of the year to account for seasonal effects on pricing.
+  \item $\text{Old Price}_i$ is a continuous independent variable representing the price of the product in the previous month, allowing for analysis of price evolution.
+  \item $\text{Vendor}_i$ is a categorical variable with levels for each vendor included in the study, capturing vendor-specific pricing strategies.
+  \item $\beta_0$ is the intercept, $\beta_0 \sim \text{Normal}(0, 2.5)$.
+  \item $\beta_1$, the coefficient for Month, $\beta_1 \sim \text{Normal}(0, 2.5)$.
+  \item $\beta_2$, the coefficient for Old Price, $\beta_2 \sim \text{Normal}(0, 2.5)$.
+  \item $\beta_{\text{vendor}}$, coefficients for each vendor level, $\beta_{\text{vendor}} \sim \text{Normal}(0, 2.5)$ for each vendor level.
+  \item $\epsilon_i$ is the error term, assumed to be normally distributed.
+\end{itemize}
 
 ```
+We conducted the Bayesian analysis in R [@citeR], utilizing the `rstanarm` package [@rstanarm].
+To maintain consistency and reliability in our modeling approach, we employed the default priors provided by `rstanarm`.
+These priors are well-suited for a wide range of data types and ensure robustness in the inference process, particularly in complex models such as ours that involve multiple predictors and hierarchical structures.
+
+Normal priors with a mean of 0 and a standard deviation of 2.5 are assigned to the coefficients and the intercept.
+These priors were chosen to impose a slight regularization effect, reducing the risk of overfitting by moderating the influence of extreme values or outliers in the data.
+
+## Model justification
+The Bayesian linear regression model was selected for this analysis to explore how various factors such as month, old price, and vendor influence the current price of organic products.
+This model type was chosen due to its flexibility in incorporating prior knowledge and handling uncertainty in estimates, which is particularly beneficial in markets where data may exhibit variability and non-standard distributions.
+Bayesian methods allow for the integration of prior beliefs or empirical evidence into the analysis, which can be particularly useful when dealing with organic product pricing where past market trends can inform current expectations.
+The primary assumptions include the linearity between predictors and the outcome, independent and identically distributed residuals, and normally distributed error terms.
+These assumptions are typical for linear regression but must be validated through diagnostic checks to confirm no significant deviations occur.
+
+Unlike frequentist regression, Bayesian regression provides a probabilistic approach to inference, offering a full posterior distribution of the parameters, which helps in understanding the uncertainty around the estimated effects.
+This model is robust to various types of data and does not strictly require the normality assumption for residuals that a classical linear regression would.
+This is particularly beneficial given the often skewed or heterogeneous nature of pricing data.
+Each predictor's influence is quantified by respective coefficients, adjusted for vendor-specific variations and temporal trends.
+The use of normal priors for these coefficients is justified by the need to regularize estimates, thus preventing overfitting and ensuring stable predictions even with potentially collinear or sparse data.
+
+### Comparison with Alternative Models
+Although a standard multiple linear regression could model continuous outcomes effectively, it lacks the ability to incorporate prior distributions and assess uncertainty comprehensively.
+odels such as logistic regression or Poisson regression were considered; however, these models are more suited to categorical or count data, respectively.
+Given that our outcome variable, current price is continuous, a linear approach is more appropriate.
+These were also evaluated, especially for capturing any non-linear dynamics between the predictors and the current price.
+While potentially offering a closer fit to certain complex patterns, the increased model complexity could hinder interpretation and require more extensive data to validate effectively.
+
+## Model Vialidation
+Model diagnostics included checks for multicollinearity, heteroscedasticity, and normality of residuals.
+Posterior predictive checks were performed to ensure the model adequately captures the observed data patterns.
+These steps help in verifying model assumptions and the appropriateness of the model for the given data.
+The background details and diagnostics are included in [Appendix -@sec-model-details].
 
@@
 ```{r}
+#| label: tbl-modelresults
+#| tbl-cap: Summary of Bayesian Model
 #| echo: false
@@
 
-pp_check(bayesian_price_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(bayesian_price_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
 ```

### 2024-12-02 paper/paper.qmd
Update Paper Result section
More graphs added

@@
 ```
+The Bayesian regression model's outputs provide insightful revelations about the factors influencing the current prices of organic products.
+These findings are encapsulated in Table 5, which succinctly presents the estimated coefficients and their respective standard errors.
+This quantitative output, complemented by the visual representation in the corresponding plot, offers a clear perspective on the relative impacts of each predictor.
+
+The intercept, estimated at 3.53, represents the average base price of organic products when all other predictor variables are held at zero.
+This baseline offers a benchmark against which the effects of other variables are measured.The coefficient for month, which is -0.42, suggests a monthly decrease in prices.
+This trend could be indicative of seasonal adjustments, where prices may drop during certain times of the year due to increased supply from harvest periods or reduced consumer demand.
+The negative value emphasizes the importance of considering temporal factors in pricing strategies.
+The positive coefficient for old_price (0.62) underscores a significant reliance on historical pricing data.
+This relationship indicates that prices are likely to increase if they were higher in the preceding period, reflecting perhaps market confidence or ongoing trends in consumer willingness to pay.
+
+The vendor coefficients reveal distinct pricing strategies among different retailer.
+vendorVoila shows a notably higher positive coefficient of 1.1), suggesting that this vendor typically prices organic products higher than others.
+This could reflect a premium branding strategy or a targeted demographic willing to pay more for perceived quality.
+In contrast, vendorNoFrills and vendorTandT have coefficients of 0.09 and -0.02, respectively, indicating minimal deviation from the baseline prices, possibly due to a competitive pricing strategy aimed at attracting cost-conscious consumers.
 
 ```{r}
+#| label: fig-vendor
+#| fig-cap: Outliers and trends in Pricing Between Different Vendors
 #| echo: false
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
 #| warning: false
+#| message: false
+
+ggplot(organic_analysis_data, aes(x = vendor, y = current_price)) +
+  geom_boxplot() +
+  labs(title = "Current Prices by Vendor", x = "Vendor", y = "Current Price") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
+```{r}
+#| label: fig-price
+#| fig-cap: Price Trends in Different Vendor Over Months
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+ggplot(organic_analysis_data, aes(x = month, y = current_price, color = vendor)) +
+  geom_line(aes(group = interaction(vendor, month)), linewidth = 1) +  # Updated line
+  geom_line(aes(y = old_price), linetype = "dashed") +
+  facet_wrap(~vendor) +
+  labs(title = "Price Trends by Vendor and Month",
+       x = "Month",
+       y = "Price ($)",
+       color = "Vendor") +
+  theme_minimal()
+```
+
+```{r}
+#| label: fig-trends
+#| fig-cap: Difference between Old Price and Current Price of Organic product
+#| eval: true
+#| echo: false
+#| warning: false
+#| message: false
+ggplot(organic_analysis_data, aes(x = month)) +
+  geom_smooth(aes(y = current_price, colour = "Current Price"), se = FALSE) +
+  geom_smooth(aes(y = old_price, colour = "Old Price"), se = FALSE) +
+  labs(title = "Current and Old Prices Over Time",
+       x = "Month",
+       y = "Price ($)",
+       color = "Price Type") +
+  scale_color_manual(values = c("Current Price" = "blue", "Old Price" = "red")) +
+  theme_minimal()
 

### 2024-12-03 paper/paper.qmd
Updated Paper result and discussion
More result section added, and discussion section finished

@@
 
-```{r}
+```{r tbl.pos='H'}
 #| label: tbl-modelresults
@@
 
-```{r}
+```{r fig.pos='H'}
 #| label: fig-vendor
@@
 
-ggplot(organic_analysis_data, aes(x = vendor, y = current_price)) +
-  geom_boxplot() +
-  labs(title = "Current Prices by Vendor", x = "Vendor", y = "Current Price") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+organic_analysis_data$price_change = organic_analysis_data$current_price - organic_analysis_data$old_price
+
+# Scatter plot of Price Change vs. Month, color-coded by Vendor
+ggplot(organic_analysis_data, aes(x = month, y = price_change, color = vendor)) +
+  geom_point(alpha = 0.6, size = 3) +
+  geom_smooth(se = FALSE) +
+  labs(title = "Monthly Price Change by Vendor",
+       x = "Month",
+       y = "Price Change ($)",
+       color = "Vendor") +
+  theme_minimal()
 
 ```
+The @fig-vendor illustrates the changes in pricing for organic products offered by six vendors over three key months: March, June, and September.
+It reveals that most vendors have adjusted their prices moderately, with the majority of changes clustering a little below the $5 mark.
+This pattern suggests a trend where vendors generally reduce their prices slightly, potentially reflecting seasonal promotions or responses to market dynamics.
+Notably, some vendors like Metro exhibit more dramatic price reductions, indicating aggressive pricing strategies or perhaps an adjustment to overstock situations.
+This subtle yet consistent price reduction across most vendors could imply a market-wide strategy to boost sales volumes by making organic products more affordable to a broader customer base during these months.
 
-```{r}
+```{r fig.pos='H'}
 #| label: fig-price
@@
 ```
+In the @fig-price, the pricing trends of presumably organic food items across six different vendors are displayed over months.
+The graph is segmented into six panels, each corresponding to a vendor: Galleria, Metro, NoFrills, TandT, Voila, and Walmart, allowing for a direct comparison of price fluctuations across different time points within the year.
+Each vendor's pricing data is presented with a unique colour code, maintaining consistency across the panels, which aids in quick visual comparison and analysis.
 
-```{r}
+The analysis of the graph reveals distinct trends and behaviors among the vendors.
+Galleria shows a minor incremental trend in pricing, suggesting a stable market position.
+In contrast, Metro displays notable volatility, with significant price spikes evident in June.
+NoFrills exhibits very limited data variability in the graph, which could be indicative of several market behaviors or strategies.
+This minimal fluctuation in pricing might suggest that NoFrills does not have a significant inventory of organic food products, perhaps due to a strategic focus on other product lines or a customer base that is less inclined towards organic options.
+TandT's prices slightly decline over the observed months, potentially reflecting seasonal promotions or inventory adjustments.
+Voila exhibits the most variability, with prices peaking significantly in summer seasons, which may reflect dynamic pricing strategies in response to demand changes or supply issues.
+Walmart, similar to Voila, shows price fluctuations with an increasing trend towards variability as the year progresses.
+
+```{r fig.pos='H'}
 #| label: fig-trends
@@
 ```
+The @fig-trends delineates the evolution of pricing data under two categories: "Current Price" and "Old Price," over the months.
+This graphical representation enables an insightful comparison of the pricing fluctuations throughout the year for a particular product or set of products.
+The current price, represented in blue, shows a downward trajectory over the year, falling from about 8 dollars to approximately 4 dollars.
+Conversely, the old price, depicted in red, demonstrates a similar decline from March to June but maintains a level higher than the current price during this period.
+Notably, there is a sharp dip in September where it plunges to below 6 dollars before slightly recovering.
+This divergence between the current and old prices from June to September, where the current price escalates substantially while the old price sharply declines, may suggest responses to external market forces such as supply chain disruptions, variations in seasonal demand, or changes in promotional strategies that impact pricing structures.
+Particularly, the pronounced drop in the old price during September could indicate strategies such as clearance sales or pricing adjustments intended to clear existing inventory or adapt to waning demand.
+
+This overall trend where the current prices fall below the old prices could indicate a reduction in the market price of organic food.
+Several factors could contribute to this phenomenon, such as increased supply of organic products, enhanced efficiency in organic farming techniques, or perhaps a decrease in consumer demand.
+Moreover, such a price drop could also be a strategic response by retailers to attract more customers or to remain competitive in the market, especially if other factors like seasonal variations or promotional efforts play a role.
 
@@
 
-## First discussion point {#sec-first-point}
+## Overview of Study {#sec-first-point}
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+This paper investigates the pricing dynamics of organic products across various vendors over distinct months, employing a Bayesian regression model to elucidate the factors that significantly influence pricing.
+The model's findings, presented in @tbl-modelresults, highlight how both temporal elements and historical pricing data impact current pricing strategies, alongside vendor-specific behaviors.
+Through a detailed examination of changes in pricing by month and vendor, the study offers a granular look at how seasonal adjustments, market confidence, and competitive strategies shape the pricing landscape of organic products.
 
-## Second discussion point
+## Insights on Market Behaviour
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+The first significant learning from this study is the clear indication of seasonal pricing impacts.
+The negative coefficient for month suggests that prices generally decrease over time, likely due to seasonal produce availability or a strategic reduction in prices to clear inventory before new stock arrives.
+Additionally, this trend likely reflects the agricultural cycles of organic products, where increased supply during harvest periods may lead to reduced prices.
+This seasonal adjustment could be a strategic response to fluctuating consumer demand, which often wanes after peak buying seasons.
+Understanding these trends is crucial for retailers and suppliers in planning inventory and promotional strategies.
 
-## Third discussion point
+A second crucial revelation from the study pertains to the significant influence of historical prices on current pricing strategies.
+The regression model reveals this through a coefficient of 0.62 for old prices, indicating a strong positive relationship.
+This relationship suggests that vendors heavily consider past pricing when setting current prices, likely aiming to maintain price consistency and build consumer trust.
+Such a strategy may reflect market confidence, where past prices help shape consumer perceptions of value and quality, especially in markets where organic products are seen as premium goods.
 
-## Weaknesses and next steps
+### Historical Pricing as a Market Predictor
+Another critical observation from the study is the role of historical prices in shaping current pricing strategies.
+The positive relationship between past and present prices suggests a form of price inertia, where past price levels serve as a baseline for current pricing decisions.
+This phenomenon could be driven by consumer perceptions of value and quality, which once established, can allow retailers to maintain higher price points.
+This aspect of the pricing strategy is particularly relevant in the organic market, where consumers are often willing to pay a premium for products they perceive as healthier or more ethically produced.
+Retailers leveraging this insight might focus on building and maintaining a strong brand reputation that supports higher pricing.
 
-Weaknesses and next steps should also be included.
+The persistence of higher prices from one period to the next suggests a robust market memory and potentially indicates consumer tolerance for higher price points.
+This finding is highly relevant in the organic market, where branding and perceived quality are pivotal.
+Consumers often associate higher prices with superior quality, and this perception enables vendors to maintain premium pricing strategies, even amidst market fluctuations.
+This insight is particularly important for organic products, where quality and ethical sourcing are often key selling points that justify higher price tags.
+Thus, understanding the impact of historical pricing on current strategies provides vendors with crucial information for crafting effective pricing models that resonate with consumer expectations and sustain market positioning.
+
+### Vendor-Specific Pricing Strategies
+The analysis also highlighted significant differences in pricing strategies among vendors.
+For instance, the vendor Voila exhibited a notably higher coefficient of 1.1, suggesting a premium pricing strategy that targets consumers willing to pay more for perceived quality.
+In contrast vendors such as NoFrills and TandT appear to have a limited selection of organic food products, as evidenced by their minimal price variability and relatively small deviations from baseline pricing.
+This could suggest that organic products are not a significant focus in their inventory, possibly due to their target demographic or market strategy.
+Both vendors might prioritize affordability and accessibility, catering to price-conscious consumers who may not prioritize organic options.
+This limited emphasis on organic products could also reflect supply chain constraints or a strategic decision to focus on high-demand, non-organic items that align with their cost-competitive positioning in the market.
+By offering only a small range of organic products, these vendors may be targeting a niche segment of their customer base while maintaining their broader appeal as budget-friendly retailers.
+
+## Limitations of the Study
+
+While the study provides valuable insights, several limitations must be acknowledged.
+First, the reliance on historical pricing data may not fully capture real-time market dynamics or consumer behavior shifts, such as changes in consumer preferences due to economic factors or health trends.
+Additionally, the model assumes linearity and continuity in price adjustments, which may not hold in real-world scenarios where pricing decisions can be abrupt or influenced by non-continuous factors like marketing campaigns or regulatory changes.
+
+## Future Directions
+
+Moving forward, several avenues can enhance our understanding of organic product pricing.
+Future research could integrate more granular data, such as specific product types within the organic category, to ascertain if different products exhibit unique pricing behaviors.
+Additionally, incorporating consumer data, such as demographics and purchase histories, could offer deeper insights into the demand side of the pricing equation.
+Methodologically, employing non-linear models or machine learning techniques might capture more complex interactions and non-linear dependencies not accommodated by the current model.
+
+Furthermore, expanding the geographic scope to include different regions or comparing urban versus rural markets could reveal broader pricing trends and strategies.
+Lastly, longitudinal studies extending over several years would help in understanding long-term trends and the effects of macroeconomic factors on organic product pricing.
 

### 2024-12-03 scripts/05-model_data.R
Fixed Preamble
Updated preamble

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Create Bayesian Model
+# Author: Dingshuo Li
+# Date: 01 December, 2024
+# Contact: dawson.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: analysis_data is saved as parquet in data files
+# Any other information needed? N/A
 
 
 #### Workspace setup ####

### 2024-12-03 paper/paper.qmd
Paper discussion and appendix finished
Appendix A finished and fixed error in discussion

@@
 
+### Data Source
 The dataset under analysis was sourced from Jacob Filipp's Hammer Project @hammerData, which compiles extensive pricing data on organic food products sold across various retail vendors.
@@
 These datasets were sourced from the Hammer Project, containing over 12 million rows of diverse product data, including non-organic items and a variety of vendors beyond the scope of this study.
-The transaction records (raw_data) include timestamps, vendor details, and product identifiers, while the product details (product_data) provide comprehensive information such as product names, brands, pricing history, and unit measurements.
+The transaction records raw_data include timestamps, vendor details, and product identifiers, while the product details product_data provide comprehensive information such as product names, brands, pricing history, and unit measurements.
 
@@
 
-# Additional data details
+# Methodological Framework for Data Collection and Analysis in Project Hammer
+
+## Introduction
+Project Hammer is an innovative initiative designed to compile and make available a comprehensive database of historical grocery prices from top grocers’ websites across Canada.
+This appendix delves into the detailed methodologies used for data collection, sampling, and observational analysis within this project.
+It also explores the application of simulation techniques and integrates extensive literature to support the methodologies used.
+
+## Data Collection Techniques
+The primary method for data compilation in Project Hammer involved automated web scraping techniques.
+Scripts were developed to systematically collect price data from various online grocery platforms.
+This method ensured a high-efficiency collection of large volumes of data while maintaining consistency across different data points.
+
+### Web Scraping: 
+The primary data for Project Hammer was meticulously extracted from various top grocers' websites across Canada.
+Automated scripts using Python libraries like BeautifulSoup and Selenium were employed to scrape historical price data, ensuring efficient and consistent data retrieval across different data points.
+### API Utilization:
+Where available, APIs provided by grocery chains were utilized to fetch data in a structured format, enhancing the accuracy and reliability of the collected data.
+This method streamlined the process, reducing the likelihood of errors commonly associated with manual data collection methods.
+
+## Data Integrity and Validation
+Data Cleaning: Rigorous cleaning processes were applied to correct anomalies and fill missing values, ensuring high data quality for analysis.
+Validation Checks: Cross-verification with in-store price checks and secondary sources was conducted periodically to validate the data accuracy.
+Automated Checks: Scripts were run to identify missing data, outliers, and inconsistencies, which were then manually reviewed to determine the cause and necessary corrections.
+
+## Sampling Strategy
+The sampling strategy for Project Hammer was carefully designed to capture a comprehensive and representative dataset of grocery prices across Canada.
+Given the complex and varied nature of the Canadian grocery market, a stratified sampling approach was deemed most suitable for ensuring the data accurately reflected the diversity of market conditions and pricing strategies employed across different geographical areas and store types.
+Below is a detailed breakdown of each component of the sampling strategy.
+
+### Geographical Stratification
+Canada was segmented into its major regions: Atlantic Canada, Quebec, Ontario, the Prairies, Alberta, and British Columbia.
+This segmentation reflects significant regional differences, including population density, urbanization, and economic activity, which can all influence grocery pricing.
+Within each region, further stratification was employed to differentiate between urban and rural areas.
+Urban areas tend to have higher competition among grocers and potentially lower prices due to larger market sizes, whereas rural areas might experience higher prices due to fewer competitors and higher logistics costs.
+
+### Random Sampling within Strata
+To achieve a representative sample within each geographical stratum, random sampling was used.
+This method aims to reduce sampling bias that could skew the analysis, such as overrepresentation of major chains or urban centers.
+From each geographical and urban stratum, grocery stores were randomly selected.
+This included large national chains, regional chains, and independent stores to capture a broad spectrum of pricing strategies.
+Random sampling was not only applied to the selection of stores but also to the timing of data collection.
+Prices were recorded at different times—weekly, bi-weekly, and monthly—to account for any temporal variations in pricing that could result from promotional activities or seasonal changes.
+
+### Importance of the Sampling Strategy
+This rigorous sampling methodology ensured that the Project Hammer dataset was robust, minimizing the potential biases that can arise from non-random sampling or over-concentration in certain geographical areas or store types.
+By reflecting the true variance across different market conditions and consumer demographics, the dataset allows for more accurate and generalizable conclusions to be drawn about pricing strategies and market dynamics in the Canadian grocery sector.
+
+## Simulation Techniques
+Simulation techniques were critical in validating the web scraping algorithms and assessing potential sampling biases for Project Hammer.
+By generating synthetic datasets that emulate the dynamics of the Canadian grocery market, especially focusing on organic food pricing, the project team rigorously tested and refined our data collection and analysis methodologies.
+
+The primary goal of these simulations was to evaluate the performance and accuracy of scraping algorithms and observe the impact of sampling strategies in a controlled environment, avoiding the costs and time associated with real data collection.
+The synthetic datasets included a wide range of organic food items with varied price points and historical price trends, accurately reflecting real market conditions.
+This allowed the team to enhance the scraping tools and sampling methods, ensuring they were robust and representative of actual market dynamics.
+
+## Linkages to Literature
+The methodologies discussed in this appendix are grounded in established research, drawing on foundational resources like the International Handbook of Survey Methodology @survey and the article Statistical Data Integration in Survey Sampling: A Review.
+The International Handbook of Survey Methodology provides guidance on survey design, implementation, and ethical considerations.
+It emphasizes the importance of aligning sampling strategies, such as stratified and cluster sampling, with the survey's objectives to ensure representative data collection.
+This aligns with the section on sampling strategies in this appendix, where these methods are tailored for accuracy and efficiency.
+Additionally, the handbook's discussion of data collection methods—such as face-to-face interviews and online surveys—connects with the appendix’s focus on selecting techniques based on the target population and study requirements.
+Furthermore, the handbook highlights ethical practices, including informed consent and confidentiality, which are reflected in this appendix.
+
+The article Statistical Data Integration in Survey Sampling: A Review @datareview examines combining different data sources to improve survey estimation and analysis.
+It addresses challenges such as inconsistent structures and biases, which relates to the section on integrating survey and observational data, where merging datasets enhances analysis.
+The article discusses statistical techniques like propensity score matching and multiple imputation to improve data validity, complementing the appendix’s emphasis on validation.
+It also highlights the use of simulation studies to test and refine sampling and data collection methods, which aligns with the simulation section in this appendix.
+
+## Conclusion
+This appendix has provided a detailed account of the sophisticated methodologies employed in Project Hammer to compile a robust database of grocery prices.
+By leveraging advanced web technologies, adhering to rigorous sampling strategies, and ensuring thorough validation processes, Project Hammer aims to foster more competition and reduce potential collusion in the Canadian grocery sector.
+These approaches not only ensure the reliability of the findings but also contribute significantly to the broader academic and legal discourse on market competition and regulatory compliance in the grocery sector.
 

### 2024-12-04 other/datasheet/datasheet_template.qmd
Upload for Datasheet
wrote datasheet and created references.bib for the datasheet

---
title: "Datasheets for Datasets: Enhancing Transparency and Accountability in Data Collection '"
subtitle: "A Comprehensive Examination of the Project Hammer Grocery Dataset"
author: 
  - Dingshuo Li
thanks: "Code and data are available at: https://github.com/dawsonlll/Organic-Product-Pricing-Analysis.git"
date: today
date-format: long
abstract: "This datasheet documents the Project Hammer Grocery Dataset, a comprehensive compilation of historical grocery prices from Canadian retailers, with a focus on organic products. The dataset was created to address gaps in publicly available structured data, enabling the analysis of price trends, vendor strategies, and seasonal variations. By capturing detailed pricing information across vendors and regions, the dataset provides insights into market dynamics and competitive behavior in Canada’s grocery sector. This resource is critical for fostering transparency, informing policy decisions, and supporting research on consumer economics and anti-competitive practices."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets. And the data came from @hammerData

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to enable a detailed analysis of grocery price trends across Canada, focusing specifically on organic products. It fills a gap in publicly available datasets by providing a structured format for analyzing price evolution and vendor-specific pricing strategies.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset was assembled by Jacob Filipp as part of Project Hammer, an initiative dedicated to promoting competition and reducing collusion in Canada's grocery sector.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The creation of the dataset was independently funded, with no specific grants associated.
4. *Any other comments?*
    - This dataset serves as a valuable resource for examining market dynamics, pricing models, and economic behaviors within the Canadian grocery industry.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- The dataset comprises records of grocery prices, including fields such as product id, product name, old price, current price, vendor, and date.
2. *How many instances are there in total (of each type, if appropriate)?*
	- It includes data from eight vendors: Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart, and Save-On-Foods, covering dates from February 28, 2024, current time.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset is designed to cover all available instances within a specified timeframe, providing a comprehensive view of the market during that period.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each entry contains raw price data, supplier details and time information, allowing analysis of price changes over time.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- The primary target variable of interest is the 'current_price' of each product.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Some price are missing becuase the data is not recorded. Any missing values were addressed through data cleaning and imputation processes to ensure completeness.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- The dataset explicitly captures relationships between old and current prices, as well as vendor-specific pricing strategies.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- Not really. But for predictive modeling, a common approach is to split the data into training (70%), validation (15%), and testing (15%) sets.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- Potential errors from web scraping were mitigated through validation checks and data cleaning procedures.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset is self-contained, with data sourced directly from publicly accessible grocery websites.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	-  The dataset does not include confidential information.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	-  The data is factual and does not contain offensive content.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- Sub-populations are identified by vendor, facilitating analysis of different pricing strategies.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No individuals can be identified from the dataset.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- The dataset does not contain sensitive personal information.
16. *Any other comments?*
	- No

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- Data was collected through web scraping and API utilization from top grocers' websites.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Automated scripts and available APIs were employed for data collection, with validation through cross-referencing and in-store price checks.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- A stratified sampling approach was used to ensure geographic and vendor diversity.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Data collection was conducted by the project's creator, Jacob Filipp.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- Data collection spanned from February 28, 2024 and up-to-date
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- The data collection process adhered to ethical guidelines, focusing on publicly available information.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- Data was obtained directly from publicly accessible websites.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- Not applicable, as data was collected from public sources.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Regular audits ensured compliance with data protection standards.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- N/A
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- The dataset was analyzed for potential impact on data subjects. As the dataset does not contain personal, sensitive or identifiable information, the risk to individual privacy or data protection is negligible. The analysis focused on ensuring compliance with ethical guidelines for the use of publicly available data, confirming that the content of the dataset meets legal standards and does not infringe intellectual property rights. 
12. *Any other comments?*
	- No

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Data underwent cleaning to correct anomalies and impute missing values, ensuring high quality for analysis.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- The raw data is archived and available at https://jacobfilipp.com/hammer/
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	-  Preprocessing was performed using R scripts, with code available  at https://github.com/dawsonlll/Organic-Product-Pricing-Analysis.git
4. *Any other comments?*
	- No

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- The dataset has been utilized for analyzing price trends and vendor pricing strategies.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- https://jacobfilipp.com/hammer/ The dataset is available for download in CSV and SQLite formats.
3. *What (other) tasks could the dataset be used for?*
	-  The dataset can be used for economic analysis, market competition studies, and legal assessments of pricing practices.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- Should be cautious of potential biases due to the data collection methods and ensure appropriate analytical techniques are applied.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- The dataset is not intended for analyses requiring personal or sensitive information.
6. *Any other comments?*
	- No

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- The dataset is publicly available for download.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	-  It is distributed in CSV and SQLite formats. No DOI found.
3. *When will the dataset be distributed?*
	- The dataset was made available on February 28, 2024
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- The dataset is provided under MIT, encouraging academic and legal use.
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- There are no known restrictions imposed by third parties.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No export controls or regulatory restrictions apply.
7. *Any other comments?*
	- No

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	-  The dataset is maintained by Jacob Filipp.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- linkedin.com/in/jacobfilipp or jacobfilipp.com 
3. *Is there an erratum? If so, please provide a link or other access point.*
	- There is no erratum currently available. Any errors identified in the dataset will be documented and corrected in subsequent updates, with notifications provided on the GitHub repository.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- The dataset will be updated periodically to include new price records, correct labeling errors, and remove obsolete entries. Updates are managed by the Project Hammer team and communicated through the GitHub repository and the official website.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- The dataset does not relate to personal data or individuals; therefore, no retention limits are applicable.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Older versions of the dataset will be archived and remain accessible for reproducibility and transparency.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- A mechanism for community contributions is available through GitHub. Contributors can propose extensions or augmentations to the dataset via pull requests, which will be reviewed and validated by the Project Hammer team before inclusion. Approved contributions will be documented in the repository’s change log.
8. *Any other comments?*
	- The dataset serves as a vital tool for academic and industry research. Its ongoing maintenance and open-source nature aim to ensure its relevance and usability over time.

\newpage

# References

### 2024-12-04 scripts/00-simulate_data.R
Fixed Scripts 00-simulate
error fixed

@@ -17,7 +17,7 @@ set.seed(304)
 #### Simulate data ####
 n <- 10000  # Number of rows to simulate
 vendors <- c("Metro", "Walmart", "NoFrills", "TandT")
-products <- c("Organic", "Non-Organic")
+products <- c("Organic")
 months <- 1:12
 
 

### 2024-12-04 scripts/02-clean_data.R
Fixed clean data process
Bug fix

@@ -45,7 +45,5 @@ cleaned_data <- joined_data %>%
   tidyr::drop_na()
 
 #### Save data ####
-write_csv(joined_data, "data/02-analysis_data/joined_data.csv")
-write_csv(cleaned_data, "data/02-analysis_data/analysis_data.csv")
 write_parquet(x = cleaned_data, sink = "data/02-analysis_data/analysis_data.parquet")
 

### 2024-12-04 scripts/02-clean_data.R
Fixed Preamble
A minor typo in the preamble

@@ -2,9 +2,9 @@
 # Purpose: Cleans the raw groceries data
 # Author: Dingshuo Li
 # Date: 25 Nov 2024 
-# Contact: dawson.lir@utoronto.ca
+# Contact: dawson.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The raw data is downloaded and is in folder
+# Pre-requisites: The raw data is downloaded and is in raw data files.
 # Any other information needed? N/A
 
 #### Workspace setup ####

### 2024-12-04 scripts/03-test_analysis_data.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests for cleaned data
+# Author: Dingshuo Li
+# Date: 03 December 2024 
+# Contact: dawson.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Cleaned data is saved in analysis data files.
+# Any other information needed? N/A
 
 
 #### Workspace setup ####
@@ -13,9 +13,10 @@ library(tidyverse)
 library(dplyr)
 library(testthat)
 library(readr)
+library(arrow)
 setwd(here::here())
 
-analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
+analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 
 
 #### Test data ####

### 2024-12-04 scripts/04-exploratory_data_analysis.R

@@ -4,7 +4,7 @@
 # Date: 27 November, 2024
 # Contact: dawson.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: analysis data is in the folder
+# Pre-requisites: Analysis data is in the analysis data file.
 # Any other information needed? N/A
 
 

### 2024-12-04 scripts/05-model_data.R
Update modelling
Fixed typo in the organic-analysis-data

@@ -13,12 +13,12 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
+organic_analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 
 ### Model data ####
 bayesian_price_model <- stan_glm(
   formula = current_price ~ month + old_price + vendor,  # Defines the regression model
-  data = analysis_data,                                 # Specifies the dataset
+  data = organic_analysis_data,                                 # Specifies the dataset
   family = gaussian(),                                  # Assumes a normal distribution for the response variable
   prior = normal(location = 0, scale = 2.5, autoscale = TRUE), # Sets normal priors for coefficients
   prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE), # Sets normal prior for the intercept

### 2024-12-04 paper/paper.qmd
Update paper
paper fixed grammar and prose

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper investigates pricing patterns in the organic food market by analyzing how vendor strategies and seasonal trends influence product prices.
+Using a dataset of organic food prices across multiple vendors and months, we employed a Bayesian regression model to understand the interplay between historical prices, vendor-specific strategies, and seasonal factors.
+The analysis revealed significant seasonal price fluctuations and demonstrated that vendors rely heavily on historical prices to inform current pricing strategies.
+These findings provide valuable ideas into consumer behaviour, vendor decision-making, and the economic dynamics of organic markets, highlighting the broader implications for market competition and sustainable agricultural practices."
 format:
@@
 
-The organic food market has witnessed a remarkable surge in demand, driven by increasing consumer awareness of health and environmental impacts.
-This demand has not only heightened the market value of organic products but has also introduced complexities into their pricing strategies across various retail platforms.
-Notably, organic food typically incurs higher production and certification costs compared to conventional food products, influencing its retail pricing and accessibility.
-This study aims to dissect these pricing strategies by examining how the prices of organic foods vary across different vendors and over successive months.
+The organic food market has experienced a significant increase in demand, driven by growing consumer awareness of health and environmental impacts.
+This demand has not only raised the market value of organic products but has also introduced challenges to their pricing strategies across various retail platforms.
+Organic foods generally incur higher production and certification costs compared to conventional products, which affects their retail pricing and accessibility.
+This study examines these pricing strategies by analyzing how organic food prices vary across vendors and over time.
 
-This paper explores the intricate dynamics of organic food pricing within a well-defined dataset that records prices from multiple vendors over a specified period.
-The analysis is structured to address a gap in existing research: the detailed quantification of how prices for organic products are structured and vary in a real-market scenario.
-Utilizing advanced Bayesian statistical methods, this research assesses the relationship between current and historical prices while considering the influence of temporal factors captured through monthly variations.
+This paper investigates the dynamics of organic food pricing using a dataset that records prices from multiple vendors over a specific period.
+The analysis addresses the need for detailed quantification of how prices for organic products are structured and vary in real-market conditions.
+By applying Bayesian statistical methods, this research evaluates the relationship between current and historical prices while considering the effect of temporal factors, such as monthly variations.
 
-Estimand: The primary estimand of this study is the price elasticity of organic foods across different market vendors and months.
-Specifically, the analysis seeks to estimate how changes in the historical prices (old prices) influence the current pricing strategies, adjusting for variations across months and between vendors.
-This will allow us to quantify the direct and interactive effects of time (month) and vendor on the current prices of organic products, providing insights into the responsiveness of organic food prices to market dynamics and vendor-specific factors.
+Estimand: The primary focus of this study is the price elasticity of organic foods across different vendors and time periods.
+Specifically, the analysis estimates how changes in old prices influence current pricing strategies, accounting for differences across months and vendors.
+This approach allows for the measurement of the direct and interactive effects of month and vendor on the current prices of organic products, offering a deeper understanding of the responsiveness of organic food prices to market conditions and vendor-specific factors.
 
-The findings reveal distinct pricing patterns that differ by vendor and demonstrate notable fluctuations corresponding to seasonal trends.
-These insights are crucial for consumers making informed purchasing decisions and for vendors strategizing to capture or expand market share in the competitive organic food sector.
-The importance of understanding these pricing mechanisms lies in its potential to influence marketing strategies, consumer spending, and ultimately, the economic sustainability of organic farming practices.
+The findings highlight distinct pricing patterns among vendors and show significant fluctuations aligned with seasonal trends.
+These results are relevant for consumers making informed purchasing decisions and for vendors optimizing strategies to maintain or grow their market share in the competitive organic food sector.
+Understanding these pricing mechanisms has implications for marketing, consumer behavior, and the economic sustainability of organic farming practices.
 
 The paper is structured to provide a clear, logical flow from methodology to implications @sec-data outlines our dataset selection and cleaning, ensuring transparency.
-The @sec-model details the Bayesian model and its rationale, directly linked to our data preparation.
-The @sec-result displays our findings with visual aids, while the @sec-discussion interprets these findings in the context of broader market dynamics.
-Finally, @sec-limitation  highlights areas for further study, acknowledging any constraints encountered.
+The @sec-model details the Bayesian model and its justifications.
+The @sec-result displays our findings with visual graphs, while the @sec-discussion interprets these findings in the context of market dynamics.
+As well as highlights areas for further study, and the limitation of study.
 
@@
 ### Data Source
-The dataset under analysis was sourced from Jacob Filipp's Hammer Project @hammerData, which compiles extensive pricing data on organic food products sold across various retail vendors.
-This dataset is particularly valuable for its comprehensive coverage of both temporal and vendor-specific pricing strategies, offering a unique lens through which seasonal trends and competitive behaviors in the organic food market can be observed.
-The choice of this dataset was motivated by its granularity and the quality of its data, which includes multiple vendors and a wide temporal range, making it ideal for a nuanced analysis of pricing dynamics.
+The dataset analyzed in this study was sourced from Jacob Filipp's Hammer Project @hammerData, which compiles detailed pricing data on organic food products across various retail vendors.
+Its extensive coverage of vendor-specific and temporal pricing strategies makes it particularly suitable for observing seasonal trends and competitive behaviors in the organic food market.
+The dataset’s granularity and quality, including data from multiple vendors over an extended period, provide a robust foundation for analyzing pricing dynamics.
 
@@
 
-The dataset used in this study comes from Jacob Filipp's Hammer Project, which provides detailed records of organic food pricing across retailers and months.
-This data collection effort captures an important economic and social phenomenon: fluctuating pricing strategies in the competitive organic food market.
-Prices are recorded monthly to reflect the dynamic nature of these market conditions, providing a comprehensive snapshot that facilitates analysis of trends and pricing behavior over time.
+The dataset analyzed in this study provides detailed records of organic food prices across vendors and months, capturing the dynamic pricing strategies within the competitive organic food market.
+Monthly price recordings reflect changing market conditions and facilitate the analysis of trends and pricing behavior over time.
 
-The primary step in capturing the phenomenon involves systematically collecting price data directly from retail sources.
-This data is gathered from grocery merchants.
-Each price tag observed represents a specific decision by a vendor regarding how much to charge for an organic product at a given time.
-So the important variables: current price which is the price of an organic product at the time of data collection.
-It is observed directly from the vendor.
-And old price which is the previously recorded price for the same product, which provides a historical comparison point.
-It helps in understanding how prices are adjusted over time.
+The data collection involved systematically recording prices directly from grocery merchants.
+Each price entry represents a vendor’s decision on how much to charge for an organic product at a given time.
+Two key variables are examined: the current price, observed directly at the time of collection, and the old price, the previously recorded price for the same product.
+The old price serves as a historical reference, aiding in the analysis of price adjustments over time.
 
-Each recorded price becomes an entry in the dataset, associated with metadata that includes the vendor, the time of collection, and the product details.
-This structured approach ensures that each entry in the dataset is a reliable and accurate representation of the real-world pricing phenomena observed.
-These entries are then ready for analytical processes that can derive meaningful insights about organic food pricing dynamics.
+Each entry in the dataset includes metadata detailing the vendor, collection date, and product information, ensuring accuracy and reliability.
+This structured approach supports robust analysis and allows for meaningful conclusions about organic food pricing dynamics.
 
 ## Data Cleanning
-The analysis begins with the consolidation of raw data from two distinct sources: product transaction records and product details.
-These datasets were sourced from the Hammer Project, containing over 12 million rows of diverse product data, including non-organic items and a variety of vendors beyond the scope of this study.
-The transaction records raw_data include timestamps, vendor details, and product identifiers, while the product details product_data provide comprehensive information such as product names, brands, pricing history, and unit measurements.
+The analysis began with the consolidation of raw data from two sources: product records and product details.
+These datasets contained over 13 million rows of diverse product data, including non-organic items and vendors outside the scope of this study.
+The product records raw_data included timestamps, vendor details, and product identifiers, while the product details dataset product_data provided information such as product names, brands, pricing history, and unit measurements.
 
-To conduct analysis, the two datasets were merged using an inner join on the product_id, which serves as a unique identifier linking transaction records to specific product details.
-This joining process ensured that each transaction record was enriched with relevant product information, creating a comprehensive view of each sale's context.
+To prepare the data, the two datasets were merged using an inner join on the product_id, a unique identifier linking transaction records to specific product details.
+This process enriched each transaction record with relevant product information, creating a detailed view of each product.
 
-Post-merger, the dataset still contained an extensive range of products and vendors, many of which were irrelevant to the focus of this study—organic food pricing dynamics among major vendors.
-To refine the dataset first did vendor selection.
-The dataset was filtered to include only transactions from major retailers known for their significant market presence and influence in organic food sales, specifically "Metro," "NoFrills," "Walmart," "Voila," "TandT," and "Galleria." This selection was based on the objective to analyze pricing strategies within the most competitive segments of the retail market.
-A further filtration was applied to isolate organic products.
-This was achieved by searching for the keyword "organic" within the product_name field, using a case-insensitive search to ensure comprehensive inclusion of all relevant items.
-This step significantly reduced the dataset size and increased its relevance to the study’s aims.
+After merging, the dataset still included a wide range of products and vendors, many of which were irrelevant to the study’s focus on organic food pricing among major vendors.
+The first step in refining the dataset was vendor selection.
+Transactions were filtered to include only those from major vendors with a significant presence in organic food sales: "Metro," "NoFrills," "Walmart," "Voila," "TandT," and "Galleria." This selection focused the analysis on pricing strategies within competitive segments of the retail market.
 
-The current_price and old_price were parsed to ensure that they were represented as numerical values, free from any non-numeric characters that could interfere with quantitative analysis.
-This parsing was crucial for accurate statistical calculations and comparisons.
+The current price and old price were parsed to ensure that they were represented as numerical values, free from any non-numeric characters that could interfere with quantitative analysis.
+This parsing was important for accurate statistical calculations and comparisons.
 The nowtime field, representing the timestamp of each transaction, was used to extract the month component.
 This transformation was essential for analyzing seasonal trends and month-to-month pricing variations.
-Lastly, entries with missing or zero values in current_price were removed to maintain the integrity of pricing analysis.
+Lastly, entries with missing or zero values in current price were removed to maintain the integrity of pricing analysis.
 Additionally, any remaining rows with incomplete data were dropped, ensuring that the dataset used in subsequent analyses was entirely robust and clean.
@@
 ### Predictor variables
-The old price variable, representing the product's price from the preceding month, serves as a critical metric for understanding the evolution of market trends and pricing strategies over time.
+The old price variable, representing the product's price from the preceding month, serves as a essential metric for understanding the evolution of market trends and pricing strategies over time.
 The summary @tbl-old-price reveal a broad range of "Old Price" from a minimum of \$1.69 to a maximum of \$38.56, with a median price of \$5.49, suggesting a median market position.
@@
 #| label: tbl-old-price
-#| tbl-cap: Summary of old price
+#| tbl-cap: Overview of Historical Price Trends
 #| echo: false
@@
 ```
-The vendor variable acts as a categorical identifier for retailers within the dataset, providing crucial insights into the diversity of pricing strategies employed across different retail environments.
-This variable is essential for comparing and contrasting market behaviours as it delineates the dataset into distinct groups based on the retailer.
-The distribution of data @tbl-Vendor shows that vendors varies significantly, with Metro being the most frequently occurring vendor with 1,981 entries, followed by Voila with 1,513 entries.
-Walmart, Galleria, TandT, and NoFrills appear with decreasing frequency, suggesting a wide range in the scale and possibly the geographic reach of these retailers.
-This spread indicates that Metro and Voila dominate this dataset, which could imply a greater influence or market share in the organic products sector compared to the other vendors.
-Understanding the pricing strategies of these vendors is key to analyzing market trends, as each vendor may react differently to market pressures, consumer preferences, and seasonal changes.
-Analyzing this variable can reveal patterns that are specific to certain types of retailers and provide valuable insights into the competitive landscape of the organic food market.
+The vendor variable serves as a categorical identifier for vendors in the dataset, enabling the analysis of pricing strategies across different retail environments.
+It is essential for comparing and contrasting market behaviors, as it divides the dataset into distinct groups based on the retailer.
+The distribution of data @tbl-Vendor shows significant variation among vendors, with Metro being the most frequently represented, with 1,981 entries, followed by Voila with 1,513 entries.
+Walmart, Galleria, TandT, and NoFrills appear less frequently, reflecting differences in the scale and possibly the geographic reach of these vendors
+
+This distribution suggests that Metro and Voila dominate the dataset, potentially indicating a larger influence or market share in the organic products sector compared to other vendors.
+Understanding the pricing strategies of these vendors is important for analyzing market trends, as each vendor may respond differently to market pressures, consumer preferences, and seasonal changes.
+Analyzing this variable helps identify patterns specific to certain types of vendors and provides a better understanding of the competitive dynamics of the organic food market.
 ```{r}
 #| label: tbl-Vendor
-#| tbl-cap: Preview of Breakdown of Organic Product Listings by Vendor
+#| tbl-cap: Distribution of Organic Product Listings Across Vendors
 #| echo: false
@@
 ```
-The "Month" variable in the dataset, recorded as a numerical value, serves as a crucial indicator for analyzing seasonal impacts on organic food pricing.
+The "Month" variable in the dataset, recorded as a numerical value, serves as a important indicator for analyzing seasonal impacts on organic food pricing.
 It reflects when the data was collected, revealing essential patterns of how prices fluctuate through different months, likely influenced by consumer demand, supply variations, and promotional periods.
@@
 In contrast, the notably lower entries in spring and winter indicate reduced activity, which could be due to offseason market slowness or less frequent data collection.
-This variable is integral to understanding how temporal factors influence pricing strategies, providing valuable insights for vendors, consumers, and analysts studying market dynamics in the organic food sector.
+This variable is integral to understanding how temporal factors influence pricing strategies, providing valuable strategy for vendors, consumers, and analysts studying market dynamics in the organic food sector.
 ```{r}
 #| label: tbl-month
-#| tbl-cap: Preview of Breakdown of Organic Product Listings by Month
+#| tbl-cap: Monthly Distribution of Organic Product Listings
 #| echo: false
@@
 The third quartile at 7.49 shows that 75% of the products fall below this price point, highlighting a pricing structure that might cater more to a mid-range consumer segment.
-This comprehensive look at current price helps in understanding the economic landscape of organic food pricing and its variability, which is crucial for analyzing market strategies and consumer behaviour.
+This look at current price helps in understanding the economic landscape of organic food pricing and its variability, which is important for analyzing market strategies and consumer behaviour.
 ```{r}
 #| label: tbl-current-price
-#| tbl-cap: Preview of Summary of Current Price
+#| tbl-cap: Summary of Current Price Distribution
 #| echo: false
@@
 
-# Model
+# Model {#sec-model}
 ## Model Overview
-In our study, we are particularly interested in understanding how the current price of organic products is influenced by several key factors: the price of the product in the previous month (old price), the vendor from which the product is purchased, and the time of year (month).
+In our study, we are particularly interested in understanding how the current price of organic products is influenced by several key factors: the price of the product in the previous month, the vendor from which the product is purchased, and the time of year.
 To systematically examine these relationships, we developed a Bayesian regression model.
@@
 We conducted the Bayesian analysis in R [@citeR], utilizing the `rstanarm` package [@rstanarm].
-To maintain consistency and reliability in our modeling approach, we employed the default priors provided by `rstanarm`.
+To maintain consistency and reliability in our modeling approach, we employed the default priors.
 These priors are well-suited for a wide range of data types and ensure robustness in the inference process, particularly in complex models such as ours that involve multiple predictors and hierarchical structures.
-
 Normal priors with a mean of 0 and a standard deviation of 2.5 are assigned to the coefficients and the intercept.
@@
 
-Unlike frequentist regression, Bayesian regression provides a probabilistic approach to inference, offering a full posterior distribution of the parameters, which helps in understanding the uncertainty around the estimated effects.
+Unlike other regression, Bayesian regression provides a probabilistic approach to inference, offering a full posterior distribution of the parameters, which helps in understanding the uncertainty around the estimated effects.
 This model is robust to various types of data and does not strictly require the normality assumption for residuals that a classical linear regression would.
@@
 ### Comparison with Alternative Models
-Although a standard multiple linear regression could model continuous outcomes effectively, it lacks the ability to incorporate prior distributions and assess uncertainty comprehensively.
+Although a standard multiple linear regression could model continuous outcomes effectively, it lacks the ability to incorporate prior distributions and assess uncertainty.
 odels such as logistic regression or Poisson regression were considered; however, these models are more suited to categorical or count data, respectively.
@@
 
-# Results
-
-Our results are summarized in @tbl-modelresults.
-
+# Results {#sec-result}
 ```{r tbl.pos='H'}
 #| label: tbl-modelresults
-#| tbl-cap: Summary of Bayesian Model
+#| tbl-cap: Overview of Bayesian Model Analysis
 #| echo: false
@@
 ```
-The Bayesian regression model's outputs provide insightful revelations about the factors influencing the current prices of organic products.
+Our results are summarized in @tbl-modelresults.
+The Bayesian regression model's outputs provide implications about the factors influencing the current prices of organic products.
 These findings are encapsulated in Table 5, which succinctly presents the estimated coefficients and their respective standard errors.
-This quantitative output, complemented by the visual representation in the corresponding plot, offers a clear perspective on the relative impacts of each predictor.
+This quantitative output, complemented by the visual representation in the corresponding plot, provides a clear perspective on the relative impacts of each predictor.
 
 The intercept, estimated at 3.53, represents the average base price of organic products when all other predictor variables are held at zero.
-This baseline offers a benchmark against which the effects of other variables are measured.The coefficient for month, which is -0.42, suggests a monthly decrease in prices.
+This baseline provide a benchmark against which the effects of other variables are measured.The coefficient for month, which is -0.42, suggests a monthly decrease in prices.
 This trend could be indicative of seasonal adjustments, where prices may drop during certain times of the year due to increased supply from harvest periods or reduced consumer demand.
@@
 The vendor coefficients reveal distinct pricing strategies among different retailer.
-vendorVoila shows a notably higher positive coefficient of 1.1), suggesting that this vendor typically prices organic products higher than others.
+vendor Voila shows a notably higher positive coefficient of 1.1, suggesting that this vendor typically prices organic products higher than others.
 This could reflect a premium branding strategy or a targeted demographic willing to pay more for perceived quality.
-In contrast, vendorNoFrills and vendorTandT have coefficients of 0.09 and -0.02, respectively, indicating minimal deviation from the baseline prices, possibly due to a competitive pricing strategy aimed at attracting cost-conscious consumers.
+In contrast, NoFrills and TandT have coefficients of 0.09 and -0.02, respectively, indicating minimal deviation from the baseline prices, possibly due to a competitive pricing strategy aimed at attracting cost-conscious consumers.
 
@@
 #| label: fig-vendor
-#| fig-cap: Outliers and trends in Pricing Between Different Vendors
+#| fig-cap: Vendor-Specific Monthly Price Variations
 #| echo: false
@@
 organic_analysis_data$price_change = organic_analysis_data$current_price - organic_analysis_data$old_price
-
-# Scatter plot of Price Change vs. Month, color-coded by Vendor
 ggplot(organic_analysis_data, aes(x = month, y = price_change, color = vendor)) +
@@
 The @fig-vendor illustrates the changes in pricing for organic products offered by six vendors over three key months: March, June, and September.
-It reveals that most vendors have adjusted their prices moderately, with the majority of changes clustering a little below the $5 mark.
+It shows that most vendors have adjusted their prices moderately, with the majority of changes clustering a little below the $5 mark.
 This pattern suggests a trend where vendors generally reduce their prices slightly, potentially reflecting seasonal promotions or responses to market dynamics.
@@
 #| label: fig-price
-#| fig-cap: Price Trends in Different Vendor Over Months
+#| fig-cap: Monthly Price Trends Across Vendors
 #| echo: false
@@
 
-The analysis of the graph reveals distinct trends and behaviors among the vendors.
+The analysis of the graph shows distinct trends and behaviors among the vendors.
 Galleria shows a minor incremental trend in pricing, suggesting a stable market position.
@@
 #| label: fig-trends
-#| fig-cap: Difference between Old Price and Current Price of Organic product
+#| fig-cap: Comparison of Old and Current Prices for Organic Products
 #| eval: true
@@
 The @fig-trends delineates the evolution of pricing data under two categories: "Current Price" and "Old Price," over the months.
-This graphical representation enables an insightful comparison of the pricing fluctuations throughout the year for a particular product or set of products.
+This graphical representation enables an comparison of the pricing fluctuations throughout the year for a particular product or set of products.
 The current price, represented in blue, shows a downward trajectory over the year, falling from about 8 dollars to approximately 4 dollars.
@@
 Several factors could contribute to this phenomenon, such as increased supply of organic products, enhanced efficiency in organic farming techniques, or perhaps a decrease in consumer demand.
-Moreover, such a price drop could also be a strategic response by retailers to attract more customers or to remain competitive in the market, especially if other factors like seasonal variations or promotional efforts play a role.
+Moreover, such a price drop could also be a strategic response by vendors to attract more customers or to remain competitive in the market, especially if other factors like seasonal variations or promotional efforts play a role.
 
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 The model's findings, presented in @tbl-modelresults, highlight how both temporal elements and historical pricing data impact current pricing strategies, alongside vendor-specific behaviors.
-Through a detailed examination of changes in pricing by month and vendor, the study offers a granular look at how seasonal adjustments, market confidence, and competitive strategies shape the pricing landscape of organic products.
+Through a detailed examination of changes in pricing by month and vendor, the study provide a granular look at how seasonal adjustments, market confidence, and competitive strategies shape the pricing landscape of organic products.
 
-## Insights on Market Behaviour
+## Market Behaviour
 
@@
 This seasonal adjustment could be a strategic response to fluctuating consumer demand, which often wanes after peak buying seasons.
-Understanding these trends is crucial for retailers and suppliers in planning inventory and promotional strategies.
+Understanding these trends is important for vendors and suppliers in planning inventory and promotional strategies.
 
-A second crucial revelation from the study pertains to the significant influence of historical prices on current pricing strategies.
-The regression model reveals this through a coefficient of 0.62 for old prices, indicating a strong positive relationship.
+A second important revelation from the study pertains to the significant influence of historical prices on current pricing strategies.
+The regression model indicated this through a coefficient of 0.62 for old prices, indicating a strong positive relationship.
 This relationship suggests that vendors heavily consider past pricing when setting current prices, likely aiming to maintain price consistency and build consumer trust.
@@
 ### Historical Pricing as a Market Predictor
-Another critical observation from the study is the role of historical prices in shaping current pricing strategies.
+Another essential observation from the study is the role of historical prices in shaping current pricing strategies.
 The positive relationship between past and present prices suggests a form of price inertia, where past price levels serve as a baseline for current pricing decisions.
-This phenomenon could be driven by consumer perceptions of value and quality, which once established, can allow retailers to maintain higher price points.
+This phenomenon could be driven by consumer perceptions of value and quality, which once established, can allow vendors to maintain higher price points.
 This aspect of the pricing strategy is particularly relevant in the organic market, where consumers are often willing to pay a premium for products they perceive as healthier or more ethically produced.
-Retailers leveraging this insight might focus on building and maintaining a strong brand reputation that supports higher pricing.
+vendors leveraging this might focus on building and maintaining a strong brand reputation that supports higher pricing.
 
@@
 Consumers often associate higher prices with superior quality, and this perception enables vendors to maintain premium pricing strategies, even amidst market fluctuations.
-This insight is particularly important for organic products, where quality and ethical sourcing are often key selling points that justify higher price tags.
-Thus, understanding the impact of historical pricing on current strategies provides vendors with crucial information for crafting effective pricing models that resonate with consumer expectations and sustain market positioning.
+This is particularly important for organic products, where quality and ethical sourcing are often key selling points that justify higher price tags.
+Thus, understanding the impact of historical pricing on current strategies provides vendors with information for crafting effective pricing models that resonate with consumer expectations and sustain market positioning.
 
@@
 This limited emphasis on organic products could also reflect supply chain constraints or a strategic decision to focus on high-demand, non-organic items that align with their cost-competitive positioning in the market.
-By offering only a small range of organic products, these vendors may be targeting a niche segment of their customer base while maintaining their broader appeal as budget-friendly retailers.
+By offering only a small range of organic products, these vendors may be targeting a niche segment of their customer base while maintaining their broader appeal as budget-friendly vendors.
 
@@
 
-While the study provides valuable insights, several limitations must be acknowledged.
+While the study provides valuable study, several limitations must be acknowledged.
 First, the reliance on historical pricing data may not fully capture real-time market dynamics or consumer behavior shifts, such as changes in consumer preferences due to economic factors or health trends.
@@
 Future research could integrate more granular data, such as specific product types within the organic category, to ascertain if different products exhibit unique pricing behaviors.
-Additionally, incorporating consumer data, such as demographics and purchase histories, could offer deeper insights into the demand side of the pricing equation.
+Additionally, incorporating consumer data, such as demographics and purchase histories, could offer deeper into the demand side of the pricing equation.
 Methodologically, employing non-linear models or machine learning techniques might capture more complex interactions and non-linear dependencies not accommodated by the current model.
@@
 ## Introduction
-Project Hammer is an innovative initiative designed to compile and make available a comprehensive database of historical grocery prices from top grocers’ websites across Canada.
+Project Hammer is an innovative initiative designed to compile and make available database of historical grocery prices from top grocers’ websites across Canada.
 This appendix delves into the detailed methodologies used for data collection, sampling, and observational analysis within this project.
@@
 ## Sampling Strategy
-The sampling strategy for Project Hammer was carefully designed to capture a comprehensive and representative dataset of grocery prices across Canada.
+The sampling strategy for Project Hammer was carefully designed to capture dataset of grocery prices across Canada.
 Given the complex and varied nature of the Canadian grocery market, a stratified sampling approach was deemed most suitable for ensuring the data accurately reflected the diversity of market conditions and pricing strategies employed across different geographical areas and store types.
@@
 ## Simulation Techniques
-Simulation techniques were critical in validating the web scraping algorithms and assessing potential sampling biases for Project Hammer.
+Simulation techniques were essential in validating the web scraping algorithms and assessing potential sampling biases for Project Hammer.
 By generating synthetic datasets that emulate the dynamics of the Canadian grocery market, especially focusing on organic food pricing, the project team rigorously tested and refined our data collection and analysis methodologies.
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check to assess the fit of the model.
+The dark blue line represents the observed data distribution, while the lighter blue lines show the simulated data distributions generated from the posterior predictive distribution of the model.
+The simulated data predicted the overall trend and the alignment between the observed data and the model-generated data indicates that the model is capturing the key characteristics of the observed data well.
+This suggests that the model is a good fit for the dataset and effectively reflects the underlying generative process.
+In @fig-ppcheckandposteriorvsprior-2 the comparison between posterior and prior distributions demonstrates how the observed data influenced the model's parameter estimates.
+The posterior distributions are narrower and more concentrated than the diffuse priors, indicating that the data provided substantial evidence to refine the estimates.
+Parameters like old_price and month show strong and precise posterior effects, highlighting their significant roles in explaining current prices.
+Vendor-specific coefficients vary, with Volia showing a notable positive effect, reflecting higher pricing strategies, while TandT and  NoFrills remain close to zero, indicating minimal deviation from baseline pricing.
+This comparison underscores how the model relies on data to inform impactful predictors while maintaining prior regularization for less informative parameters, offering thinking into vendor-specific strategies and seasonal pricing trends.
 
@@
 
+# Posterior predictive check
+pp_check(bayesian_price_model) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+
+# Posterior vs prior comparison
+posterior_vs_prior(bayesian_price_model) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  coord_flip()
+
 ```
@@
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+@fig-stanareyouokay-1 is a trace plot.
+It illustrates the convergence and mixing of the Markov Chain Monte Carlo (MCMC) sampling process for the Bayesian model parameters.
+Each parameter's plot Intercept, month, old_price and vendors are shown.
+The chains show consistent oscillations around a stable range without any noticeable trends or drifts, indicating that the sampler has likely converged.
+Parameters like old_price and vendorVoila display tightly clustered chains, suggesting precise estimates with relatively low uncertainty.
+In contrast, parameters such as vendorNoFrills and vendorTandT exhibit greater variability, reflecting higher uncertainty or minimal influence on the model.
+Overall, the chains demonstrate good overlap and mixing, which is important for thoroughly exploring the posterior distributions.
+These observations suggest that the MCMC algorithm has successfully converged, and the parameter estimates from the model are robust and reliable.
 
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+@fig-stanareyouokay-2 is a Rhat plot.
+It shows the R-hat diagnostic for all parameters in the Bayesian model.
+The R-hat statistic assesses the convergence of Markov Chain Monte Carlo simulations, comparing the variance within chains to the variance between chains.
+In this plot, all the R-hat values are at or very close to 1.00, well below the threshold of 1.1, indicating that the chains for all parameters have converged.
+The R-hat values being near 1.00 signify excellent mixing of the chains, meaning that each chain explores the parameter space effectively and independently of its starting values.
 

### 2024-12-04 other/datasheet/datasheet_ Hammer.qmd
Update paper and datesheet
Fixed typo

--- 2024-12-04 other/datasheet/datasheet_template.qmd
@@
 ---
-title: "Datasheets for Datasets: Enhancing Transparency and Accountability in Data Collection '"
+title: "Datasheets for Datasets: Enhancing Transparency and Accountability in Data Collection"
 subtitle: "A Comprehensive Examination of the Project Hammer Grocery Dataset"

### 2024-12-04 paper/paper.qmd
Update paper and datesheet
Fixed typo

@@
 All data analysis was conducted using R @citeR, including statistical computing and graphics.
-And the following R packages were used for conduct code in scripts: tidyverse @tidy, Palmerpenguins @palmerpenguins, ggplot2 @ggplot2, Dplyr @citedplyr, Knitr @citeknitr, modelsummary @modelsummary, arrow @arrow, here @here, rstanarm @rstanarm, sclaes @scales, splines @splines,broom @broom, lubridate @date, janitor @jan and testthat @testthat.
+And the following R packages were used for conduct code in scripts: tidyverse @tidy, Palmerpenguins @palmerpenguins, ggplot2 @ggplot2, Dplyr @citedplyr, Knitr @citeknitr, modelsummary @modelsummary, arrow @arrow, here @here, rstanarm @rstanarm, sclaes @scales, splines @splines,broom @broom, lubridate @date, janitor @jan, stringr @stringr and testthat @testthat.
 Following  @tellingstories we conducted data simulate, test simulated data, data cleanning, test analysis data, EDA, data modelling.

### 2024-12-15 other/datasheet/datasheet_Hammer.qmd
Update the datasheet
fixed some error in datasheet

--- 2024-12-04 other/datasheet/datasheet_ Hammer.qmd
@@
 If so, please provide a description.*
-	-  The dataset does not include confidential information.
+	-  No, the dataset does not contain any confidential or personally identifiable information. All data is publicly available and pertains to grocery prices listed on vendor websites.
 12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
-	-  The data is factual and does not contain offensive content.
+	- No, the dataset contains neutral information about grocery prices, which is unlikely to cause offense or anxiety.
 13.
@@
 If so, please describe how.*
-	- No individuals can be identified from the dataset.
+	- No, the dataset contains no personal identifiers or data that could be used to identify individuals.
 15.
@@
 16. *Any other comments?*
-	- No
+	- The dataset provides a valuable resource for analyzing pricing strategies and market trends in the grocery sector.
+It does not involve data from vulerable populations, does not contain any personally identifiable information, and no formal ethics review was conducted as the dataset does not involve sensitve or private data.
 
@@
 If so, please describe how.*
-	- Data was collected through web scraping and API utilization from top grocers' websites.
+	- The data was acquired through web scraping and API utilization from the websites of major grocery vendors.
+This data includes directly observable pricing and product details.
+Validation was performed by cross-referencing online prices with in-store checks and ensuring consistency across multiple data collection runs.
 2.
@@
 7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
-	- Data was obtained directly from publicly accessible websites.
+	- The data was collected directly from publicly accessible grocery vendor websites.
 8.
@@
 If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
-	- Not applicable, as data was collected from public sources.
+	- Not applicable, as the data was collected from publicly available online sources without involving individuals.
 9.
@@
 If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
-	- Regular audits ensured compliance with data protection standards.
+	- Not applicable, as the dataset does not involve personal data or data from individuals. The focus is exclusively on publicly available product and pricing information.
 10.
@@
 If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
-	- N/A
+	- Not applicable, as no individual-level data requiring consent was collected.
 11.
@@
 12. *Any other comments?*
-	- No
+	- This dataset is a valuable tool for understanding pricing trends and consumer behaviour in the grocery market.
+However, users should note that this dataset focuses on publicly accessible data and has inherent limitations in representing small or independent retailers.
 
@@
 If not, you may skip the remaining questions in this section.*
-	- Data underwent cleaning to correct anomalies and impute missing values, ensuring high quality for analysis.
+	- Yes, the data underwent preprocessing and cleaning to address anomalies and missing values.
+Steps included correcting outliers, imputing missing values, and ensuring consistency in product names and vendor labels.
+The cleaned data was prepared to support accurate and meaningful analysis
 2.
@@
 7. *Any other comments?*
-	- No
+	- While efforts will be made to maintain the dataset, its long-term availability depends on continued project support and hosting arrangements
 
